---
title: 'Problem Set 4: Randomized Control Trials'
author: "Claire Duquennois"
output:
 html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Name: Kiersten Kochanowski**

# Empirical Analysis using Data from Bryan, G., Chowdury, S., Mobarak, A. M. (2014, Econometrica)
This exercise uses data from Bryan,Chowdhury, and Mobarak's paper, "Underinvestment in a Profitable Technology: the Case of Seasonal Migration in Bangladesh," published in *Econometrica* in 2014. This paper studies the effects of seasonal migration on household consumption during the lean season in rural Bangladesh by randomly subsidizing the cost of seasonal migration. 

# Set Up: 

```{r}
#Packages
library(haven)
library(dplyr)
library(stargazer)
library(lfe)
```


## Finding the data
The data can be found by going to Mushfiq Mobarak's Yale faculty page, select "data", and then following the link to the data repository page on the Harvard dataverse. You will need to sign in to get access to the data files. Once logged in, you will find many possible files to download. Navigate to the second page of listed files and download `Mobarak - Monga Dataverse files.zip` which contains all the files we need. 

```{r}
#Kiersten's WD
getwd()
Round1_Ctrl <- read_dta("/Users/Kiersten/Documents/School/S2_Causal_Inf/PS4_Kochanowski/Data/Round1_Controls_Table1.dta")
Round2 <- read_dta("/Users/Kiersten/Documents/School/S2_Causal_Inf/PS4_Kochanowski/Data/Round2.dta")
Round1_Ctrl <- as.data.frame(Round1_Ctrl)
View(Round1_Ctrl)
Round2 <- as.data.frame(Round2)
View(Round2)
```

## Question 1: Loading the data - Load any packages you will need and the data contained in the following files `Round1_Controls_Table1.dta` and `Round2.dta`. How many observations are contained in each of these datasets. What is the level of an observation? Explain any discrepancies between the datasets.**

**Code:**
```{r}
nrow(Round1_Ctrl)
nrow(Round2)
```

**Answer:**
There are 1900 observations in the Control Table 1 data frame and 1907 observations in the Round 2 data frame. When the NGO did the round 1 interviews it was in October of 2008 and at the time some of the household individuals were in the city and therefore omitted in the round 1 data, and those missing individuals were recovered in the round 2 data.

The level of the observations is the household (HHID).
 
## Question 2: Data Description- The dataset contains many variables, some of which are not used in this exercise. Keep the following variables in the final datasets (Hint: use the `select` function in `dplyr`).** 

For Round 1 data:

| Name            |Description                                                             |
|-----------------|------------------------------------------------------------------------|
|cash             |In cash treatment group                                                 |
|credit           |In credit treatment group                                               |
|info             |In information treatment group                                          |
|control          |In control group                                                        |
|q9pdcalq9        |Total calories per person per day                                       | 
|exp_total_pc_r1  |Total monthly household expenditures per capita                         |
|hhmembers_r1     |Number of household members                                             |
|tsaving_hh_r1    |Total household savings                                                  |
|hhh_education    |Household head is educated                                              |
|num_adltmalesr1  |Number of adult males in the household                                  |

Code
```{r}
Round1_f <- select(Round1_Ctrl, cash, credit, info, control, q9pdcalq9, exp_total_pc_r1, hhmembers_r1, tsaving_hh_r1, hhh_education, num_adltmalesr1)
View(Round1_f)
```


For Round 2 data:

| Name          |Description                                                             |
|---------------|------------------------------------------------------------------------|
|cash           |In cash treatment group                                                 |
|credit         |In credit treatment group                                               |
|info           |In information treatment group                                          |
|control        |In control group                                                        |
|average_exp2   |Total consumption per person per month in round 2                       |
|lit            |Highest reading and writing ability of household                        |
|walls_good     |Wall material (income proxy)                                            |
|monga          |Subjective expectations about monga at baseline                         |
|dhaka_remit    |Subjective expectations about migration remitances at baseline          |
|dhaka_network  |Subjective expectations about social network in city at baseline        |
|exp_total_pc_r1|Total household expenditures per capita at baseline                     |
|subsistencer1  |Share of food out of total expenditures at baseline                     |
|num_adltmalesr1|Household adult males at baseline                                       |
|num_childrenr1 |Household small children at baseline                                    |
|avgQ13earned   |Average skill score of network                                          |
|constrainedr1  |Denied or ineligible for credit at baseline                             |
|bankedr1       |Has received credit at baseline                                         |
|upazila        |Sub-district name                                                       |
|village        |Village name                                                            |
|migrant        |Member of household migrates this season                                |
|total_fish     |Total monthly household expenditures per capita on fish                 |
|migrant_new    |Household has a first time migrant this season                          |

```{r}
Round2_f <- select(Round2, cash, credit, info, control,average_exp2, lit, walls_good, monga, dhaka_remit, dhaka_network, exp_total_pc_r1, subsistencer1, num_adltmalesr1, num_childrenr1, avgQ13earned, constrainedr1, bankedr1, upazila, village, migrant, total_fish, migrant_new)
View(Round2_f)
```

# Analysis:

## **Question 3: Regress all the baseline household characteristics still included in the round 1 data on the following three variables: $cash_i$, $credit_i$ and $info_i$, and present your results in a table. What is the equivalent table in the paper?** 

**Code:**
```{r, Q3, results ='asis'}
ra <- felm(data=Round1_f, cash ~ hhmembers_r1 + hhh_education + num_adltmalesr1 + q9pdcalq9 + exp_total_pc_r1 + tsaving_hh_r1)
rb <- felm(data=Round1_f, credit ~ hhmembers_r1 + hhh_education + num_adltmalesr1 + q9pdcalq9 + exp_total_pc_r1 + tsaving_hh_r1)
rc <- felm(data=Round1_f, info ~ hhmembers_r1 + hhh_education + num_adltmalesr1 + q9pdcalq9 + exp_total_pc_r1 + tsaving_hh_r1)
rd <- felm(data=Round1_f, control ~ hhmembers_r1 + hhh_education + num_adltmalesr1 + q9pdcalq9 + exp_total_pc_r1 + tsaving_hh_r1)

stargazer(ra, rb, rc, rd, type="html", header=FALSE, title="Randomization Balance on Observables at Baseline", 
                     style="qje", 
                     se = list(ra$rse, rb$rse, rc$rse, rd$rse),
                     column.labels = c("Cash Group", "Credit Group", "Info Group", "Control Group"),
                     omit = "Constant",
                     covariate.labels = c("Household Size", "Head of House Educ.", "Number of Adult Males", "Total Cal/Person/Day", "Monthly Exp/Person", "Household Savings"))
```

**Answer:** 
This is a balance table which reveals whether there are any biases in the groups (control and experimental) at baseline. It mirrors Table 1 in the paper.

## **Question 4: How should the coefficients in the table above be interpreted? What should we look for in this table?**

**Answer:**
We are looking to see whether there are statistically significant relationships between the observable pre-treatment characteristics and the groups households were assigned to. If any of these were significant, it could indicate that the assignment of groups was not random (or that it is coincidence per statistical chance). In our case, none of the coefficients are significant which means assignment was random and balanced across groups.

## **Question 5: Using the round 2 data, regress migrant on the treatment arm indicators. What is the equivalent table in the paper?**

**Code:**
```{r, Q5, results ='asis'}
r4 <- felm(data=Round2_f, cash ~ migrant)
r5 <- felm(data=Round2_f, credit ~ migrant)
r6 <- felm(data=Round2_f, info ~ migrant)
r7 <- felm(data=Round2_f, control ~ migrant)

stargazer(r4, r5, r6, r7, type="html", header=FALSE, title="Program Participation (Migration) Rates", 
                     style="qje", 
                     se = list(r4$rse, r5$rse, r6$rse, r7$rse),
                     covariate.labels = c("Migrant"),
                     dep.var.labels.include = TRUE,
                     omit = "Constant",
                     column.labels = c("Incentivized", "Not Incentivized"),
                     column.separate = c(2,2))
```

**Answer:** 
This mirrors Table 2 in the paper titled "Program Up-Take Rates"

## **Question 6: How should the coefficients in the table above be interpreted? Why is this table important?**

**Answer:**
This table is important in order to determine whether incentives had an impact on households' decisions to migrate during the monga, or whether they're in line with the control group's migration rate.

## **Question 7: What is the underlying migration rate in the control group and how might this change our interpretation of the results? **

**Answer:**
Although my table's results aren't accurate, the migration rate in the control group is ~36% which means when we look at the effect of treatment, we must look at the change in migration rate with and without treatment relative to a baseline of 36%.


## **Question 8: Replicate the results presented in the third row of the first three columns of table 3. Present them in a table and interpret these results.**

**Code:**
```{r}
#Cluster package
library(cluster)
```


```{r}
#Drop outlier
max(Round2_f$total_fish)
Round2_f_sub <- Round2_f[-c(92),]
max(Round2_f_sub$total_fish)

```

```{r}
#Model
r8<-felm(average_exp2~cash+credit+info|upazila|0|village, data=Round2_f_sub)
summary(r8)
```

```{r, Q8, results ='asis'}
#Save cluster robust standard errors
cluster_se8 <- as.vector(summary(r8,cluster = c("village"))$coefficients[,"Cluster s.e."])

stargazer(r8, type="html", header=FALSE, title="Effects of Migration on Consumption Among Remaining Household Members", 
                     style="qje", 
                     se = list(r8$cluster_se8),
                     covariate.labels = c("Cash", "Credit", "Info"),
                     dep.var.labels.include = FALSE,
                     omit = "Constant",
                     column.labels = c("Total consumption"))
```


**Answer:**
This table shows that incentive treatments (cash and credit), which induced more migration during monga, resulted in significant increases in household consumption among remaining members in the villages. Households in the cash treatment group spent 96.57 more taka/capita on average (at the 99% confidence level), households in the credit treatment group spent 76.74 more taka/capita on average (at the 95% confidence level). The caveat with these results is that we'd expect consumption per capita to increase given there's one less person in the household during monga (the migrant).


## **Question 9: What happens to these estimates if you drop the fixed effects from the specification. Why? **

**Code:**
```{r}
r9<-felm(average_exp2~cash+credit+info|0|0|village, data=Round2_f_sub)
summary(r9)

cluster_se9 <- as.vector(summary(r9,cluster = c("village"))$coefficients[,"Cluster s.e."])
```

```{r, Q9, results ='asis'}
stargazer(r8, r9, type="html", header=FALSE, title="Effects of Migration on Consumption Among Remaining Household Members in 2008", 
                     style="qje", 
                     se = list(r8$cluster_se8, r9$cluster_se9),
                     covariate.labels = c("Cash", "Credit", "Info"),
                     dep.var.labels.include = FALSE,
                     omit = "Constant",
                     add.lines = list(c("Sub-District FE", "Yes", "No")),
                     column.labels = c("Total consumption"))
```

**Answer:**
When missing the sub-district control, our treatment coefficients are downwardly biased.


## **Question 10: Replicate the results presented in the third row of the fourth and fifth columns of table 3. What happens to the coefficient and standard errors? Is this surprising? What does this tell us?**

Hint: You will need to construct a new variable to run these estimates. 

**Code:**
```{r}
#Constructing ITT variable
Round2_f_sub$ITT<-0
Round2_f_sub$ITT[Round2_f_sub$cash=="1"]<-1
Round2_f_sub$ITT[Round2_f_sub$credit=="1"]<-1

#Model with ITT
r10 <- felm(average_exp2~ITT|upazila|0|village, data=Round2_f_sub)
cluster_se10 <- as.vector(summary(r10,cluster = c("village"))$coefficients[,"Cluster s.e."])

#Model with ITT and following controls: household education, proxy for income, pctg of total expenditure on food, number of adult males, number of children, lacked access to credit, borrowing, total household expenditures/capita at baseline, expectations of support
r11 <- felm(average_exp2 ~ ITT + lit + walls_good + subsistencer1 + num_adltmalesr1 + num_childrenr1 + bankedr1 + exp_total_pc_r1 + monga + dhaka_network + constrainedr1 + bankedr1|upazila|0|village, data=Round2_f_sub)
cluster_se11 <- as.vector(summary(r11,cluster = c("village"))$coefficients[,"Cluster s.e."])
```


Display results
```{r, Q10, results='asis'}
stargazer(r10, r11, type="html", header=FALSE, title="Effects of Migration on Consumption Among Remaining Household Members", 
                     style="qje", 
                     se = list(r10$cluster_se10, r11$cluster_se11),
                     covariate.labels = c("Intent to Treat"),
                     dep.var.labels.include = FALSE,
                     omit = c("Constant", "lit", "walls_good", "subsistencer1", "num_adltmalesr1", "num_childrenr1", "bankedr1", "exp_total_pc_r1", "monga", "dhaka_network", "constrainedr1"),
                     column.labels = c("Total consumption", "Total consumption"),
                     add.lines = list(c("Additional controls", "No", "Yes")))
```

**Answer:**
Given that the significance and magnitude of the ITT coefficient doesn't change much when controlling for baseline characteristics, the results/relationship are robust between ITT and total consumption.

## **Question 11: Why is the header of the first five columns of table 3 "ITT". What is meant by this and what does this tell us about how we should interpret these results?**

**Answer:**
The header represents the relationship between households assigned to some treatment and their food consumption (excludes households in the control group). ITT stands for "intent to treat" which has a different meaning than the effect of being treated. Instead, its the effect of being assigned to the treatment group which means some households may not comply (and their actions are accounted for in the average ITT effect). 


## **Question 12: We are interested in estimating how migration affects total expenditures for the households that were induced to migrate by the cash and credit treatments as follows,**

$$
TotExp_{ivj}=\alpha+\beta_1Migrate_{ivj}+\theta X_{ivj}+\varphi_j+\nu_{ivj}
$$
**where $Migrate_{ivj}$ is dummy indicator for if a member of household i in village v in subdistrict j migrated, $X_{ivj}$ is a vector of control variables and $\varphi_j$ are the subdistrict fixed effects. However it is not possible to identify in the data which households were induced by the treatment vs those who would have migrated  either way. Furthermore, there is likely substantial selection between the households that select into migration versus those that do not. Propose a source of exogenous variation that can be used as an instrument to isolate "good" exogenous variation in migration. **

**Answer:**
We know the assignment of treatment was random given the controlled experiment. This ITT variable can be a source of good, exogenous variation on the action of migrating (the true action/treatment we're trying to capture).
 

## **Question 13: What is the first stage specification?**

**Answer:**
We must determine whether there's a strong relationship between ITT and migration.
$$
Migrant_{ivj}=\alpha+\beta_1ITT_{ivj}+\nu_{ivj}+\epsilon_{ivj}
$$

Where migrant is a binary variable, beta is the indicator for whether or not household was assigned to treatment (incentive) and v represents baseline controls.

## **Question 14: Estimate the first stage and check that you have a strong instrument for migration.**

Note: The first stage results reported in the paper appendix may differ slightly as explained in the table footnote.  

**Code:**
```{r, Q14, results='asis'}

#Checking strong first stage
r12 <- felm(data=Round2_f_sub, migrant ~ ITT)
r13 <- felm(data=Round2_f_sub, migrant ~ cash)
r14 <- felm(data=Round2_f_sub, migrant ~ credit)

#Displaying results
stargazer(r12, r13, r14, type="html", header=FALSE, title="Relationship between Treatment Assignment & Migration", 
                     style="qje", 
                     se = list(r12$rse, r13$rse, r14$rse),
                     covariate.labels = c("Intent to treat", "Cash", "Credit", "Info"),
                     dep.var.labels.include = FALSE,
                     omit = "Constant",
                     column.labels = c("Migration", "Migration", "Migration"))
```

**Answer:**
The first stages are strong for each of the IVs (ITT, cash, and credit all against their relationship to migration) given the statistical significance. The magnitude of the coefficients is also helpful when we rely on the exogenous variation's impact on independent variable migration (and ultimately its impact on total consumption).


## **Question 15: Use your instrument to estimate the LATE (Local Average Treatment Effect), the impact of migration on total consumption for those induced to migrate by the treatment, as in columns 6 and 7 of table 3 in the paper. Interpret your results. **

Note: if you wish to replicate the paper's coefficients exactly, you will need to use multiple instruments, one for each treatment arm.

**Code:**
```{r, Q15a}
r15 <- felm(data=Round2_f_sub, average_exp2 ~ 1 |upazila|(migrant~ITT)|village) 
r16 <- felm(data=Round2_f_sub, average_exp2 ~ walls_good + subsistencer1 + num_adltmalesr1 + num_childrenr1 + bankedr1 + exp_total_pc_r1 + monga + dhaka_network + constrainedr1 + bankedr1|upazila|(migrant~ITT + walls_good + subsistencer1 + num_adltmalesr1 + num_childrenr1 + bankedr1 + exp_total_pc_r1 + monga + dhaka_network + constrainedr1 + bankedr1)|village)

```

```{r, Q15b, results='asis'}
stargazer(r15, r16, type="html", header=FALSE, title="Local Average Treatment Effect of Migration on Total Consumption", 
                     style="qje", 
                     se = list(r15$rse, r16$rse),
                     dep.var.labels.include = TRUE,
                     omit = c("upazila", "Constant", "walls_good", "subsistencer1", "num_adltmalesr1", "num_childrenr1", "bankedr1", "exp_total_pc_r1", "monga", "dhaka_network", "constrainedr1", "bankedr1"),
                     covariate.labels = c("IV"),
                     add.lines = list(c("Controls", "No", "Yes")),
                     column.labels = c("Total Consumption", "Total Consumption"))
```

**Answer:**
These regressions show the effect of migration (only for those households that were induced to migrate by the RCT's intervention) on consumption. The coefficient for the model without any controls shows that there was a 374.86 increase in taka spent/capita for households that had an individual migrate (on average). When controlling for household socio-economic characteristics, this effect is still significant and robust: there was a 308.998 increase in taka spent/capita for households that had an individual migrate (on average).


## **Question 16: Why are these results different from those in columns 4 and 5 of the paper?**

**Answer:** 
The LATE results are different from the previous regressions (ITT) because they estimate the effect of treatment (migration) and not just the effect of being in the treatment group (which is susceptible to non-compliers). 

## **Question 17: Why is this value particularly relevant for policy decisions in the context of this experiment.**

**Answer:**
If policy makers were to apply the findings of this research in the real world, they'd likely be concerned with both the LATE and ITT estimators. The LATE will help policy makers determine whether migration is a worthwhile tool to help bring more economic stability to households vulnerable to poverty during monga season. The ITT estimator provides an idea of how public assistance in support of migration may impact behavior and outcomes for households, taking into account that not everyone will behave as the program would like.

## **Question 18: Suppose a policy maker found these results so compelling that they decided to make this a national policy. How would general equilibrium effects potentially change the impacts of this policy if it was implemented in a very large scale way?**

**Answer:**
If a national policy was implemented, mass migration could affect the amount of jobs available in cities (whereas with an RCT the volume of individuals migrating doesn't impact the GE) which could greatly impact the effectivity of temporary migration into cities in order to earn more money when jobs are scarce in local villages.

## **Question 19: One major concern that is often brought up in discussions about RCT's is the problem of external validity. It is not always clear how informative the findings from a small scale research project in one context are for policy makers working on a different scale and in different contexts. What are your thoughts on the external validity of this particular project and RCT's in general? **

**Answer:**
I think if this RCT was to be rolled out on a national level the same results may be more difficult to achieve (like all RCT applications) for a few reasons: NGOs may have more resources, less beauracy, and benefit from a smaller scale to manage these trials compared to government agencies (aka pilot bias)  


# Submission instructions:

1) Make sure the final version of your assignment is uploaded on GitHub in both html and Rmarkdown format. 




